### 2024-11-29 paper/paper.qmd
initial uploads

---
title: "Predicting Goodreads Ratings Across Genres"
subtitle: "My subtitle if needed"
author: 
  - Dhruv Gupta
thanks: "Code and data are available at: https://github.com/dhruv5423/GoodReads-Cross-Genre-Predictions"
date: today
date-format: long
toc: true
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

# Introduction

Goodreads, founded in 2006, boasts over 150 million users and is perhaps the most popular online community for4 readers. At it's heart, Goodreads is akin to a reader's social media platform, allowing them to display, discover, review, and catalog books they love (or hate) across multiple different genres.

Personal Motivation

Estimand

Results

Broader Context

# Data

## Sourcing

## Note on Measurement

## Outcome Variables

## Predictor Variables

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
## Model Set-Up

### Model Justification

# Results

# Discussion

## First Discussion Point

## Second Discussion Point

## Third Discussion Point

## Weaknesses and Next Steps

\newpage

# Appendix

# Addtional Data Details

# Model Details

## Posterior Predictive Check

## Diagnostics

\newpage

# References