### 2024-12-03 paper/paper.qmd
donwload data, raw data, and simulate data

---
title: "How Well Do Strike Rates Translate from T20 to ODI Formats for Top 100 vs. Non-Top 100 players?"
subtitle: "My subtitle if needed"
author: 
  - Dhruv Gupta
thanks: "Code and data are available at: https://github.com/dhruv5423/Top-100-Batters-T20-to-ODI-Strike-Rate-Mapping"
date: today
date-format: long
toc: true
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

# Introduction

Personal Motivation

Estimand

Results

Broader Context

# Data

## Sourcing

## Note on Measurement

## Outcome Variables

## Predictor Variables

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
## Model Set-Up

### Model Justification

# Results

# Discussion

## First Discussion Point

## Second Discussion Point

## Third Discussion Point

## Weaknesses and Next Steps

\newpage

# Appendix

# Addtional Data Details

# Model Details

## Posterior Predictive Check

## Diagnostics

\newpage

# References