### 2024-12-04 other/datasheet/datasheet_template.qmd
final changes to paper, sketches, datasheet,

---
title: "Datasheet for CricSheet's T20 and ODI Datasets"
author: 
  - Dhruv Gupta
date: today
date-format: long
abstract: "This datasheet contains information about Stephen Rushe's datasets and his website, CricSheet, which he has created"
format: pdf
number-sections: true
bibliography: references.bib
---

Extract of the questions from @gebru2021datasheets. The data used is from CricSheet, an open-source platform created by Stephen Rushe [@cricsheet]

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The dataset was created to support cricket analytics and research, specifically to enable detailed analysis of cricket matches. It provides granular ball-by-ball data for T20 and ODI matches, allowing for performance modeling, strategy analysis, and player comparison.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - The dataset was created by Stephen Rushe and is made available through the CricSheet platform.
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - The dataset is publicly available and maintained by Stephen Rushe on a volunteer basis, with some support from donations.
4. *Any other comments?*
    - The dataset has become a critical resource for researchers and cricket enthusiasts, given its high level of detail and accessibility.

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- Each instance represents a ball-by-ball event in a cricket match, capturing details such as the batsman, bowler, runs scored, extras, and type of dismissal.
2. *How many instances are there in total (of each type, if appropriate)?*
	- The dataset contains 17,944 cricket matches, including 2,499 T20 Internationals and 2,391 ODIs, with millions of individual ball events.

3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- The dataset is a comprehensive record of all publicly available international cricket matches but does not include all domestic or non-international games.

4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- Each instance includes "raw" ball-by-ball data, capturing attributes such as match ID, over and ball number, batting team, bowler, batsman, runs, extras, and dismissal type.

5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- No explicit labels are included, but data attributes such as runs and dismissals provide context for outcome modeling.

6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	- Data for some older matches might be incomplete, particularly regarding player details or venue information.

7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- Yes, instances are linked within a match via match ID and innings number.

8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- No explicit splits are recommended, but users often create splits by match type (ODI vs. T20) or time period.

9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- Minor inconsistencies or missing data may exist for older matches due to incomplete record-keeping.

10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- Yes, the dataset is self-contained and does not rely on external resources.

11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- No, the dataset consists of publicly available information.

12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- No, the data is focused on professional cricket and does not include sensitive personal information.

13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- Yes, it identifies players, teams, and match types.

14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- Yes, players are identifiable, but the data is already public.

15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- The dataset is designed to be open and accessible for a wide range of analytical uses.

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- The data was sourced from official match scorecards and ball-by-ball commentary.

2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- Data was collected manually and through automated scraping tools, validated against official records.

3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- The dataset includes all publicly available international matches.

4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- The dataset is curated and maintained by Stephen Rushe.

5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- The data spans decades of cricket history, with ongoing updates.

6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- Not applicable, as the data is publicly available.

7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- No, the data was compiled from official scorecards.

8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- Not applicable, as the data is public.

9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- Not applicable, as the data is public.

10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- Not applicable, as the data is public.

11. *Any other comments?*
	- The dataset's reliability depends on the accuracy of the source records.

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- Yes, the dataset has been used for various cricket-related analyses, including player performance evaluation, predictive modeling of match outcomes, comparative analysis between match formats (e.g., ODI vs. T20), and strategy planning for teams.
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- While there is no centralized repository for all works using the dataset, individual academic papers, blogs, and analyses frequently reference CricSheet. Examples include cricket analytics blogs and statistical analyses published by enthusiasts and researchers.

3. *What (other) tasks could the dataset be used for?*
	- The dataset could be used for a wide variety of cricket-related tasks, including:
    - Forecasting player and team performance.
    - Identifying trends in gameplay over time.
    - Strategy optimization for teams and players.
    - Fan engagement through visualizations and interactive tools.
    - Developing machine learning models for predictive analytics in cricket.

4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- The dataset does not contain sensitive or personal information but relies on the accuracy of publicly available data. Inaccuracies in source data could propagate into analyses. Additionally, care must be taken when applying findings to avoid perpetuating biases in player evaluation or stereotyping based on performance in specific contexts.

5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- The dataset should not be used for unethical purposes such as match-fixing or gambling-related activities. Additionally, it should not be used to make decisions that could negatively impact individuals without proper validation of findings.

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- Yes, it is openly available on the CricSheet website.

2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- Via CSV files on the CricSheet website.

3. *When will the dataset be distributed?*
	- The dataset is continually updated and is currently available.

4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- Open license, as specified by CricSheet.

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- Stephen Rushe.

2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- Contact information is available on the CricSheet website. His Mastodon server and social page can be found at https://social.deeden.co.uk/@srushe

3. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- Yes, it is updated regularly with new match data.

4 *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- Older data remains available as part of the dataset.

\newpage

# References